Hard Statistics in Economics: Investigating Economic Drivers in Individual Confidence in the DOW in the US

### 2024-12-04 paper/paper.qmd
last edits of quarto doc

In order to determine predictive performance of this model, we can look at the predicted confidence indexes as given by the VAR model compared to the observed confidence indexes we purposely omitted form the dataset. This can be seen in @tbl-predicted. Here we ask the VAR to predict the next 8 quarters of one-year confidence indexes.

```{r}
#| label: tbl-predicted
#| tbl-cap: Model summary of the predicted relation of consumer confidence in the stock market and several economic factors..
#| echo: false
#| warning: false
#| message: false

@@

All things considered, there is a still a valuable outcome to this paper and it is that "hard" economic statistics are not enough to be able to predict whether an individual feels confident that the market will increase in one year. It shows that our suspicions may have been true that any confidence or lack of confidence in the market come from other places. Perhaps it is the knowledge of smaller sectors of the market or maybe that a sort of market psychology is a better predictor. The only concrete results we can see from this paper is that unemployment has a positive effect on market confidence and that previous market confidence is the biggest predictor of its future.

## Weaknesses and next steps {#sec-limitations}

Weaknesses and next steps should also be included.

@@

# Model details {#sec-model-details}
@tbl-summary shows the summary output of the VAR model discussed above. Note that the model specific to this paper is last output of estimated coefficients where the one-year confidence index is a function of all other variables at time $t-1$.

@@

```{r}
#| label: tbl-summary
#| tbl-cap: Model summary of the predicted relation of consumer confidence in the stock market and several economic factors..
#| echo: false
#| warning: false
#| message: false
library(modelsummary)
# Read Model
# analysis_model <- readRDS(here::here("models/best_var_model.rds"))
summary(analysis_model)

>>>>

In order to determine predictive performance of this model, we can look at the predicted confidence indexes as given by the VAR model compared to the observed confidence indexes we purposely omitted form the dataset. This can be seen in @fig-predicted. Here we ask the VAR to predict the next 8 quarters of one-year confidence indexes.

```{r}
#| label: fig-predicted
#| fig-cap: Model summary of the predicted relation of consumer confidence in the stock market and several economic factors.
#| echo: false
#| warning: false
#| message: false

@@

## Takeaways
All things considered, there is a still a valuable outcome to this paper and it is that "hard" economic statistics are not enough to be able to predict whether an individual feels confident that the market will increase in one year. It shows that our suspicions may have been true that any confidence or lack of confidence in the market come from other places. Perhaps it is the knowledge of smaller sectors of the market or maybe that a sort of market psychology is a better predictor. The only concrete results we can see from this paper is that unemployment has a positive effect on market confidence and that previous market confidence is the biggest predictor of its future. This indicates that the future outlook on the market is more psychological and less responsive to economic fundamentals than we might have expected.

## Weaknesses and next steps {#sec-limitations}
Despite considerable efforts to develop a robust model, its predictive power remains limited. This limitation may stem from several factors. One potential reason is that the model attempts to predict outcomes during a period when COVID-19 was at its peak, which may have introduced unique, unaccounted-for influences on market confidence. Additionally, it is possible that the economic factors included in the model are simply not strongly related to market confidence. Another important consideration is that the model relies solely on quantitative data, overlooking the potential impact of categorical factors that could play a crucial role in predicting the one-year confidence index.

There are several next steps that could enhance this study. First, expanding the model to incorporate additional quantitative economic factors, such as industry-specific data, could provide a more comprehensive understanding of market confidence. Moreover, given our findings that market confidence may be influenced by psychological factors, incorporating elements from behavioral economics, such as investor sentiment and speculation, could improve prediction accuracy. It is also possible that the predictive challenges stem from the model's inability to capture complex, non-linear relationships. In such a case, leveraging machine learning techniques might help identify and model these intricate patterns. Finally, some of the unexpected findings, such as the positive relationship between rising unemployment and market confidence, merit further exploration to better understand the underlying causes and implications.\newpage

@@

# Model details {#sec-model-details}
@fig-summary shows the summary output of the VAR model discussed above. Note that the model specific to this paper is last output of estimated coefficients where the one-year confidence index is a function of all other variables at time $t-1$.

@@

```{r}
#| label: fig-summary
#| fig-cap: Model summary of the predicted relation of consumer confidence in the stock market and several economic factors..
#| echo: false
#| warning: false
#| message: false

summary(analysis_model)
```
